---
title: "HMO"
output: html_document
date: "2023-04-23"
---

## R Data Cleaning

```{r}

#install.packages("rworldmap", dependencies = TRUE )


library(arules)
library(kernlab)
library(caret) 
library(tidyverse)
library(ggplot2)
library(arulesViz)
library(rworldmap)


HMO=read.csv("https://intro-datascience.s3.us-east-2.amazonaws.com/HMO_data.csv",header=TRUE)
head(HMO)
```

```{r}
which(is.na(HMO$X))
```

```{r}
which(is.na(HMO$age))
```

```{r}
which(is.na(HMO$bmi))
```

```{r}
which(is.na(HMO$children))
```

```{r}
which(is.na(HMO$smoker))
```

```{r}
which(is.na(HMO$location))
```

```{r}
which(is.na(HMO$location_type))
```

```{r}
which(is.na(HMO$education_level))
```

```{r}
which(is.na(HMO$yearly_physical))
```

```{r}
which(is.na(HMO$exercise))
```

```{r}
which(is.na(HMO$married))
```

```{r}
which(is.na(HMO$hypertension))
```

```{r}
which(is.na(HMO$gender))
```

```{r}
which(is.na(HMO$cost))
```

```{r}
HMO$bmi[is.na(HMO$bmi)] = mean(HMO$bmi, na.rm = TRUE)
```

```{r}
uniqv <- unique(HMO$hypertension)
uniqv[which.max(tabulate(match(HMO$hypertension, uniqv)))]
HMO$hypertension[is.na(HMO$hypertension)]= uniqv[which.max(tabulate(match(HMO$hypertension, uniqv)))]
```

```{r}
which(is.na.data.frame(HMO))
```

```{r}
#write.csv(HMO,"~/Downloads/Cleaning.csv", row.names = FALSE)
```


```{r}
        ###### Data Modeling and Trend #########


#install.packages("dplyr")              # Install dplyr package

library(plotly) #Used plotly library
library("dplyr") #Used dplyr library


# Histogram to visual representation of count of people with active exercise vs Not-Active

hist(HMO$age,col="Yellow",main = "Age Distribution")

# Box Plot [ BMI vs Location Type ]

df_bmi_location <- HMO[, c('bmi','location_type')]   # Subset the data to BMI and Location Type

ggplot(df_bmi_location, aes(x=location_type, y=bmi)) + geom_boxplot(fill=c("blue","green")) + labs(x='Location Type', y='BMI') + ggtitle('BMI vs. Location Type') 
  
  
# Bar Chart to observe visual representation of active smokers with respect to location


library(ggplot2) # ggplot library referenced 

num_smokers_by_location <- table(HMO$location[HMO$smoker == "yes"]) # Subset the data to Smoker as 'Yes' value

df_smokers_by_location <- data.frame(location = names(num_smokers_by_location), num_smokers = as.numeric(num_smokers_by_location)) # Data frame created for Smokers and it's count by converting to numeric type 

ggplot(df_smokers_by_location, aes(x = location, y = num_smokers)) + geom_bar(stat = "identity", fill = "blue") +labs(title = "Number of smokers by location", x = "Location", y = "Number of smokers")           # Bar chart ploted and aesthetics details provided
 







```

```{r}

library(tidyverse)
# Load the data set of healthcare customers
df <- read.csv("https://intro-datascience.s3.us-east-2.amazonaws.com/HMO_data.csv")
colnames(df)

View(df)

summary(df)

colnames(df)

#check if any col has na value
sum(is.na(df$age))
sum(is.na(df$bmi))
sum(is.na(df$children))
sum(is.na(df$smoker))
sum(is.na(df$location))
sum(is.na(df$location_type))
sum(is.na(df$education_level))
sum(is.na(df$yearly_physical))
sum(is.na(df$exercise))
sum(is.na(df$married))
sum(is.na(df$hypertension))
sum(is.na(df$gender))
sum(is.na(df$cost))




#from above, bmi and hypertension has na value
#install.packages("imputeTS")
library(imputeTS)
#use interpolation to replace the nas of bmis
df$bmi <- na_interpolation(df$bmi, option = "linear")
# filter rows whre hypertention = na
df <- df[!is.na(df$hypertension),]


# double check the na count, and find both are o
sum(is.na(df$bmi))
sum(is.na(df$hypertension))


# make a copy and make sure the df_tgt is the dataframe for the project
df_tgt <- df


#install packages
#install.packages("Hmisc")
#install.packages("car")
#install.packages("corrplot")
#install.packages("neuralnet")
#install.packages("RSNNS")

# Recap of the plot
library(ggplot2)
library("Hmisc")
library("car")
library("corrplot")
library("neuralnet")
library("RSNNS")



#summary(df_tgt)
linearmodel <- lm(cost~age+bmi+children+smoker+hypertension,data = df_tgt)
summary(linearmodel)




viftable<- vif(linearmodel)
sorttable <- sort(viftable,decreasing=TRUE)
sorttable
#From VIF analysis, no significant factor found from cost to age, bmi, childer, smokeryes, hypertension


summary(df_tgt)



# age vs cost by exercise
ggplot(data=df_tgt,aes(x=age, y=cost,colour =  exercise))+ geom_point() +
  geom_smooth(method = "lm")+
  theme_dark()


# age vs cost by yearly_physical
ggplot(data=df_tgt,aes(x=age, y=cost,colour =  yearly_physical))+ geom_point() +
  geom_smooth(method = "lm")+
  theme_dark()


# bmi vs cost by yearly_physical
ggplot(data=df_tgt,aes(x=bmi, y=cost,colour =  yearly_physical))+ geom_point() +
  geom_smooth(method = "lm")+
  theme_dark()



# bmi vs cost by yearly_physical
ggplot(data=df_tgt,aes(x=bmi, y=cost,colour =  yearly_physical))+ geom_point() +
  geom_smooth(method = "lm")+
  theme_dark()


# age vs cost by  hypertension  
ggplot(data=df_tgt,aes(x=age, y=cost,colour =   hypertension ))+ geom_point() +
  geom_smooth(method = "lm")+
  theme_dark()




# bmi vs cost by exercise 
ggplot(data=df_tgt,aes(x=bmi, y=cost,colour =  exercise ))+ geom_point() +
  geom_smooth(method = "lm")+
  theme_dark()



#make copy to train df
df_t <- df_tgt


fig1 <-ggplot(data=df_t, aes(x=age, y=cost))+
  geom_bar(stat="identity",fill="lightblue")+
  theme_dark()
fig1


#install.packages("dplyr")

library(dplyr)
library(ggplot2)

age_mean_cost <- df_tgt %>%
  group_by(age) %>%
  summarise_at(vars(cost), list(name = mean)) %>%
  arrange(desc(name)) 
  
fig2 <-ggplot(data=age_mean_cost, aes(x = age, y = name))+
  geom_col() +
  theme_linedraw()
fig2


```



ltype_mean_cost <- df_tgt %>%
  group_by(gender) %>%
  summarise_at(vars(cost), list(name = mean)) %>%
  arrange(desc(name)) 
fig3 <-ggplot(data=ltype_mean_cost, aes(x = gender, y = name))+
  geom_col() +
  theme_linedraw()
fig3




library(ggplot2)
ltype_mean_cost <- df_tgt %>%
  group_by(gender) %>%
  summarise_at(vars(cost), list(name = mean)) %>%
  arrange(desc(name)) 
ggplot2::ggplot(data=ltype_mean_cost, aes(x = gender, y = name))+
  geom_col() +
  theme_linedraw()



state_mean_cost <- df_tgt %>%
  group_by(location) %>%
  summarise_at(vars(cost), list(name = mean)) %>%
  arrange(desc(name))
state_mean_cost


by_state_count <- df_tgt %>%
  count(location) %>%
arrange(desc(n))

by_state_count



#install.packages("maps")
#install.packages("mapproj")
#install.packages("reshape2")
#install.packages("kernlab")
#install.packages("caret")
#install.packages("e1071")

library(maps)
library(ggmap)
library("dplyr")

us<-map_data("state")
df_tgt$location<-tolower(df_tgt$location)
agg1<-aggregate(df_tgt$cost,by=list(df_tgt$location),FUN=mean)
agg2<-aggregate(df_tgt$cost,by=list(df_tgt$location),FUN=max)
agg3<-aggregate(df_tgt$cost,by=list(df_tgt$location),FUN=min)
agg1<-agg1%>%rename(location=Group.1)
agg2<-agg2%>%rename(location=Group.1)
aggmerge1<-merge(agg1,agg2,by = "location" )
agg3<-agg3%>%rename(location=Group.1)
aggmerge2<-merge(aggmerge1,agg3,by= "location")
aggmerge2<-aggmerge2%>%rename(min=x,average=x.x,max=x.y)
agg4<-aggmerge2[,c(2:4)]
usmerge<-merge(us,aggmerge2,all.x=TRUE,by.x="region",by.y="location")
usmerge<-usmerge%>%arrange(order)

usmap1<-ggplot(usmerge)+geom_polygon(aes(x=long,y=lat,group=group,fill=average),color="green")+coord_map()
usmap1



usmap2<-ggplot(usmerge)+geom_polygon(aes(x=long,y=lat,group=group,fill=max),color="green")+coord_map()
usmap2

usmap3<-ggplot(usmerge)+geom_polygon(aes(x=long,y=lat,group=group,fill=min),color="green")+coord_map()
usmap3


# create subset from numerical varaiables to df1
df1 <- df_t %>%
  select(age,bmi,children,cost)
  
  
# Calculate df1 correlations
df1_cor <- round(cor(df1),digits = 4)
df1_cor





  #heat map of correlations
library(reshape2)
df1_melt <- melt(df1_cor)
fig4 <- ggplot(data = df1_melt, aes(x=Var1, y=Var2,fill=value)) +geom_tile()
fig4



boxplot(df_t$cost,
  ylab = "cost",
  main = "cost boxplot"
)


#We can see for >75th quantile, there're a very long tail of cost


#study outlier 5 number
quantile(df_t$cost,probs = c(0.25,0.5,0.75,1))
  

print(mean(df_t$cost))
print(range(df_t$cost))
quantile(df_t$cost, probs = seq(from=0.75,to=1,by=0.05))


print(mean(df_t$cost))
print(range(df_t$cost))
quantile(df_t$cost, probs = seq(from=0.75,to=1,by=0.05))


cost_th <- quantile(df_t$cost,probs = c(0.75))
cost_th


#make a copy
df_new_tgt <- df_t
df_new_tgt_tgt <- df_new_tgt

str(df_t)
str(df_new_tgt)
colnames(df_t)
colnames(df_new_tgt)

# remove the cost column from df_new_tgt
#df_new_tgt <- df_new_tgt[,-14]

colnames(df_new_tgt)

head(df_new_tgt)

str(df_new_tgt)

ggplot(df_new_tgt, aes(x = bmi, fill = cost)) + 
  geom_histogram() +                                              # Draw ggplot2 histogram with manual colors
  scale_fill_manual(values = c("TRUE" = "red","FALSE" = "lightblue")) +
  theme_dark()


ggplot(df_new_tgt, aes(x = age, fill = cost)) + 
  geom_histogram() +                                              # Draw ggplot2 histogram with manual colors
  scale_fill_manual(values = c("TRUE" = "red","FALSE" = "lightblue")) +
  theme_dark()
  
  
  
ggplot(df_new_tgt, aes(x = exercise, fill = cost)) +                                            # Draw ggplot2 histogram with manual colors
  geom_histogram(stat="count")  
  
  
ggplot(df_new_tgt, aes(x = education_level, fill = cost)) +                                            # Draw ggplot2 histogram with manual colors
  geom_histogram(stat="count")  
  
 
# create training set
set.seed(110)
ratio_of_tgt = 0.8
df_new_tgt_index <- sample(nrow(df_new_tgt), ratio_of_tgt * nrow(df_new_tgt) ,replace = FALSE)
df_new_tgt_train <- df_new_tgt[df_new_tgt_index, ]
df_new_tgt_test <- df_new_tgt[-df_new_tgt_index, ]


str(df_new_tgt)


#convert df_new_tgt_test chr to factor
df_new_tgt$smoker <- as.factor(df_new_tgt$smoker)
df_new_tgt$location <- as.factor(df_new_tgt$location)
df_new_tgt$location_type <- as.factor(df_new_tgt$location_type)
df_new_tgt$education_level <- as.factor(df_new_tgt$education_level)
df_new_tgt$yearly_physical <- as.factor(df_new_tgt$yearly_physical)
df_new_tgt$exercise <- as.factor(df_new_tgt$exercise)
df_new_tgt$married <- as.factor(df_new_tgt$married)
df_new_tgt$gender <- as.factor(df_new_tgt$gender)
df_new_tgt$cost <- as.factor(df_new_tgt$cost)

head(df_new_tgt)

df_new_tgt_train$smoker <- as.factor(df_new_tgt_train$smoker)
df_new_tgt_train$location <- as.factor(df_new_tgt_train$location)
df_new_tgt_train$location_type <- as.factor(df_new_tgt_train$location_type)
df_new_tgt_train$education_level <- as.factor(df_new_tgt_train$education_level)
df_new_tgt_train$yearly_physical <- as.factor(df_new_tgt_train$yearly_physical)
df_new_tgt_train$exercise <- as.factor(df_new_tgt_train$exercise)
df_new_tgt_train$married <- as.factor(df_new_tgt_train$married)
df_new_tgt_train$gender <- as.factor(df_new_tgt_train$gender)
df_new_tgt_train$cost <- as.factor(df_new_tgt_train$cost)


str(df_new_tgt_train)

str(df_new_tgt_test)


df_new_tgt_test$smoker <- as.factor(df_new_tgt_test$smoker)
df_new_tgt_test$location <- as.factor(df_new_tgt_test$location)
df_new_tgt_test$location_type <- as.factor(df_new_tgt_test$location_type)
df_new_tgt_test$education_level <- as.factor(df_new_tgt_test$education_level)
df_new_tgt_test$yearly_physical <- as.factor(df_new_tgt_test$yearly_physical)
df_new_tgt_test$exercise <- as.factor(df_new_tgt_test$exercise)
df_new_tgt_test$married <- as.factor(df_new_tgt_test$married)
df_new_tgt_test$gender <- as.factor(df_new_tgt_test$gender)
df_new_tgt_test$cost <- as.factor(df_new_tgt_test$cost)




levels(df_new_tgt_train$location)

levels(df_new_tgt_test$location)

levels(df_new_tgt_test$location) <- levels(df_new_tgt_train$location)

str(df_new_tgt_test)

library(kernlab)
library(caret)
library(e1071)


set.seed(110)
ratio_of_tgt = 0.8
train_list <- createDataPartition(y=df_new_tgt_train$cost ,p=ratio_of_tgt,list=FALSE)
train_df <- df_new_tgt_train[train_list,]
test_df <- df_new_tgt_train[-train_list,]


sapply(train_df, function(x) length(unique(x)))
train_df[,3] <- factor(train_df[,3])


head(train_df)


#install.packages("kernlab")

library(kernlab)

svm_model <- ksvm(cost ~ ., data=train_df,C = 1,cross = 3, prob.model = TRUE)
svm_model

# training error of 11.91 %, training accuracy of 88.09 %


# make predictions for future validation
svmpredict <- predict(svm_model, test_df, type = "response")

  
#  confusion matrix with svmpredict
CM <- confusionMatrix(svmpredict, test_df$cost)
CM

summary(svmpredict)


svmpredict <- predict(svm_model, df_new_tgt_test, type = "response")
CM1 <- confusionMatrix(svmpredict1, df_new_tgt_test$cost)
CM1


#install.packages("rpart.plot")
library("rpart")
library("rpart.plot")


dfTree <- rpart(cost ~ ., data = df_new_tgt)
prp(dfTree, faclen = 0, cex = 0.8, extra = 1)
dfTree



#remove X col
assoc_data <- df_new_tgt_train[,-1]
str(assoc_data)

#install.packages("arules")
#install.packages("arulesViz")

library(arules)
library(arulesViz)




par(mar = c(0,0,0,0))

cartTree <- rpart(cost ~ ., data = df_new_tgt)
prp(cartTree, faclen = 0, cex = 0.8, extra = 1)
cartTree




predictValues <- predict(cartTree, newdata=df_new_tgt)
predictValues <- predict(cartTree, newdata=df_new_tgt, type = "class")
confMatrix <- table(predictValues,df_new_tgt$expensive)
confMatrix



accuracy <- 1 - (sum(confMatrix) - sum(diag(confMatrix)))/sum(confMatrix)
accuracy



library(caret)
# makes the sampling predictable
set.seed(110)
# Randomly sample elements to go into a training data set
trainList <- createDataPartition(y=df_new_tgt$expensive,p=.80,list=FALSE)
# Include all of those elements in the training set
trainSet <- df_new_tgt[trainList,]
# Construct test set from everything that didn’t go into the training
testSet <- df_new_tgt[-trainList,]


head(trainList)

cartTree <- rpart(expensive~ ., data = trainSet)
prp(cartTree, faclen = 0, cex = 0.8, extra = 1)



#Note use of “class”
predictValues <- predict(cartTree,newdata=testSet,type = "class")
#simpler way to do confusion matrix
confusion <- confusionMatrix(predictValues,testSet$expensive)
confusion$overall[1]




cartTree1 <- train(expensive ~ ., data=trainSet, method="rpart")
prp(cartTree1$finalModel, faclen = 0, cex = 0.8, extra = 1)


predictValues <- predict(cartTree1,newdata=testSet, type = "raw")
confusion <- confusionMatrix(predictValues, testSet$expensive)
confusion


varImp(cartTree)

varImp(cartTree1)
plot(varImp(cartTree1))



varImp(cartTree)
varImp(cartTree1)
plot(varImp(cartTree1))



